---
title: "15 Things You Didn't Know About Texas Schools"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{15 Things You Didn't Know About Texas Schools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 5,
  cache = TRUE
)
```

```{r load-packages}
library(txschooldata)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
```

Texas public schools enroll over 5.5 million students across 1,200+ districts, making it the second-largest school system in the United States. This analysis explores the most striking patterns in Texas enrollment data from 2020-2024.

```{r fetch-data, cache=TRUE}
# Fetch 5 years of enrollment data
enr <- fetch_enr_multi(2020:2024, use_cache = TRUE)
```

---

## 1. COVID Erased a Decade of Growth in One Year

The pandemic caused the largest single-year enrollment drop in Texas history. Between 2020 and 2021, Texas public schools lost **120,133 students** -- equivalent to the entire enrollment of El Paso ISD, the 7th largest district in the state.

```{r covid-drop}
state_trend <- enr %>%
  filter(is_state, subgroup == "total_enrollment", grade_level == "TOTAL") %>%
  select(end_year, n_students) %>%
  arrange(end_year) %>%
  mutate(
    change = n_students - lag(n_students),
    pct_change = round(change / lag(n_students) * 100, 2)
  )

stopifnot(nrow(state_trend) > 0)
print(state_trend)
```

```{r covid-plot}
print(state_trend)

ggplot(state_trend, aes(x = end_year, y = n_students)) +
  geom_line(linewidth = 1.2, color = "#003366") +
  geom_point(size = 3, color = "#003366") +
  geom_vline(xintercept = 2020.5, linetype = "dotted", color = "gray50") +
  annotate("text", x = 2020.5, y = 5580000, label = "COVID", color = "gray40", size = 3.5) +
  geom_segment(aes(x = 2020, xend = 2021,
                   y = state_trend$n_students[state_trend$end_year == 2020],
                   yend = state_trend$n_students[state_trend$end_year == 2021]),
               color = "#CC0000", linewidth = 1.2, linetype = "dashed") +
  annotate("text", x = 2020.5, y = 5380000, label = "-120,133",
           color = "#CC0000", fontface = "bold", size = 4) +
  scale_y_continuous(labels = comma, limits = c(5300000, 5600000)) +
  labs(
    title = "Texas Public School Enrollment",
    subtitle = "The COVID dip and slow recovery",
    x = "School Year (End)", y = "Total Students"
  ) +
  theme_minimal()
```

Enrollment has since recovered, surpassing pre-pandemic levels by 2023. But the demographic composition of that recovery tells a very different story.

---

## 2. One in Four Students is Now an English Learner

The proportion of **Limited English Proficient (LEP)** students grew by 4.1 percentage points in just five years -- from 20.3% to 24.4% of all students. This is the single largest demographic shift in the data.

```{r lep-growth}
lep_trend <- enr %>%
  filter(is_state, subgroup == "lep", grade_level == "TOTAL") %>%
  select(end_year, n_students, pct) %>%
  mutate(pct_display = round(pct * 100, 1))

stopifnot(nrow(lep_trend) > 0)
print(lep_trend %>% select(end_year, n_students, pct_display))
```

```{r lep-plot}
print(lep_trend %>% select(end_year, n_students, pct_display))

ggplot(lep_trend, aes(x = end_year, y = pct * 100)) +
  geom_line(linewidth = 1.2, color = "#E69F00") +
  geom_point(size = 3, color = "#E69F00") +
  geom_vline(xintercept = 2020.5, linetype = "dotted", color = "gray50") +
  geom_hline(yintercept = 25, linetype = "dashed", color = "gray50") +
  annotate("text", x = 2024.3, y = 25.5, label = "1 in 4", color = "gray40", size = 3.5) +
  scale_y_continuous(limits = c(19, 27)) +
  labs(
    title = "Rise of English Learners",
    subtitle = "LEP students as % of total enrollment",
    x = "School Year (End)", y = "Percent"
  ) +
  theme_minimal()
```

This has profound implications for staffing, curriculum, and resource allocation across the state. Schools need more ESL teachers, bilingual programs, and translated materials than ever before.

---

## 3. Coppell ISD is Texas's First Asian-Majority School District

In a striking demographic shift, **Coppell ISD** became Texas's first Asian-majority public school district, with **56.7%** of students identifying as Asian in 2024. Nearby **Frisco ISD** is close behind at 42.6% -- up from 33.6% in just three years.

```{r asian-districts}
# Districts with highest Asian percentage
asian_top <- enr %>%
  filter(is_district, subgroup == "asian", grade_level == "TOTAL", end_year == 2024) %>%
  inner_join(
    enr %>% filter(is_district, subgroup == "total_enrollment",
                   grade_level == "TOTAL", end_year == 2024) %>%
      select(district_id, total = n_students),
    by = "district_id"
  ) %>%
  filter(total >= 10000) %>%
  arrange(desc(pct)) %>%
  select(district_name, total, n_students, pct) %>%
  mutate(pct = round(pct * 100, 1)) %>%
  head(10)

stopifnot(nrow(asian_top) > 0)
print(asian_top)
```

```{r frisco-trend}
# Frisco's Asian population growth
frisco_asian <- enr %>%
  filter(grepl("FRISCO", district_name), is_district,
         grade_level == "TOTAL", subgroup == "asian") %>%
  select(end_year, n_students, pct) %>%
  mutate(pct = round(pct * 100, 1))

stopifnot(nrow(frisco_asian) > 0)
print(frisco_asian)

ggplot(frisco_asian, aes(x = end_year, y = pct)) +
  geom_line(linewidth = 1.2, color = "#CC79A7") +
  geom_point(size = 3, color = "#CC79A7") +
  geom_text(aes(label = paste0(pct, "%")), vjust = -0.8, size = 3.5) +
  scale_y_continuous(limits = c(30, 50)) +
  labs(
    title = "Frisco ISD: Asian Student Growth",
    subtitle = "Percentage of total enrollment",
    x = "School Year (End)", y = "Percent Asian"
  ) +
  theme_minimal()
```

These districts in the Dallas-Fort Worth metroplex reflect changing immigration and migration patterns, with significant growth in families from India, China, and other Asian countries.

---

## 4. Fort Worth ISD Lost 14% of Its Students

While statewide enrollment recovered after COVID, urban districts continue to hemorrhage students. **Fort Worth ISD** lost **11,801 students** (-14.3%), making it the fastest-declining large district in Texas.

```{r urban-decline}
# Calculate 2020-2024 changes using district_id (2020 data has NA names)
d2020 <- enr %>%
  filter(is_district, subgroup == "total_enrollment",
         grade_level == "TOTAL", end_year == 2020) %>%
  select(district_id, n_2020 = n_students)

d2024 <- enr %>%
  filter(is_district, subgroup == "total_enrollment",
         grade_level == "TOTAL", end_year == 2024) %>%
  select(district_id, district_name, n_2024 = n_students)

losses <- d2020 %>%
  inner_join(d2024, by = "district_id") %>%
  mutate(
    change = n_2024 - n_2020,
    pct_change = round((change / n_2020) * 100, 1)
  )

# Largest percentage losses among big districts
top_losses <- losses %>%
  filter(n_2020 >= 10000) %>%
  arrange(pct_change) %>%
  select(district_name, n_2020, n_2024, change, pct_change) %>%
  head(10)

stopifnot(nrow(top_losses) > 0)
print(top_losses)
```

Houston ISD's absolute loss of **25,706 students** represents more students than 90% of Texas districts even enroll.

---

## 5. IDEA Public Schools Grew 55% in Five Years

**IDEA Public Schools**, a charter network operating across Texas, grew from 49,480 to 76,819 students between 2020 and 2024 -- a gain of **27,339 students** (+55.3%). It is the fastest-growing large school system in the state.

```{r idea-growth}
# Use district_id join to include 2020 (which has NA district_name)
idea_trend <- losses %>%
  filter(grepl("IDEA", district_name))

stopifnot(nrow(idea_trend) > 0)
print(idea_trend %>% select(district_name, n_2020, n_2024, change, pct_change))

# Year-by-year for 2021-2024 (name filter works for these years)
idea_yearly <- enr %>%
  filter(district_name == "IDEA PUBLIC SCHOOLS", is_district,
         subgroup == "total_enrollment", grade_level == "TOTAL") %>%
  select(end_year, n_students) %>%
  arrange(end_year)

stopifnot(nrow(idea_yearly) > 0)
print(idea_yearly)
```

```{r charter-networks}
# Top growing large districts (2020-2024, using district_id join)
growth <- losses %>%
  filter(n_2020 >= 5000) %>%
  arrange(desc(pct_change))

top_growth <- growth %>%
  select(district_name, n_2020, n_2024, change, pct_change) %>%
  head(10)

stopifnot(nrow(top_growth) > 0)
print(top_growth)
```

Charter networks and suburban districts are absorbing students from declining urban districts, fundamentally reshaping Texas public education.

---

## 6. Fort Bend ISD: A District With No Majority

**Fort Bend ISD** is one of the most diverse school districts in America. No racial group exceeds 28% of enrollment -- White (13.2%), Black (27.8%), Hispanic (26.7%), and Asian (27.6%) students are nearly equally represented.

```{r fort-bend}
fb_demo <- enr %>%
  filter(district_name == "FORT BEND ISD", is_district, grade_level == "TOTAL",
         subgroup %in% c("white", "black", "hispanic", "asian")) %>%
  select(end_year, subgroup, pct) %>%
  mutate(pct = round(pct * 100, 1)) %>%
  pivot_wider(names_from = subgroup, values_from = pct)

stopifnot(nrow(fb_demo) > 0)
print(fb_demo)
```

```{r diversity-chart}
# Fort Bend demographics over time
fb_plot_data <- enr %>%
  filter(district_name == "FORT BEND ISD", is_district, grade_level == "TOTAL",
         subgroup %in% c("white", "black", "hispanic", "asian")) %>%
  mutate(pct = pct * 100)

stopifnot(nrow(fb_plot_data) > 0)
print(fb_plot_data %>% select(end_year, subgroup, pct))

ggplot(fb_plot_data, aes(x = end_year, y = pct, fill = subgroup)) +
  geom_area(alpha = 0.8) +
  scale_fill_manual(
    values = c("asian" = "#CC79A7", "black" = "#009E73",
               "hispanic" = "#0072B2", "white" = "#E69F00"),
    labels = c("Asian", "Black", "Hispanic", "White")
  ) +
  labs(
    title = "Fort Bend ISD: A District With No Majority",
    subtitle = "Demographic composition 2021-2024",
    x = "School Year", y = "Percent", fill = NULL
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

---

## 7. Kindergarten Enrollment Dropped 5.8%

**Kindergarten enrollment fell from 383,585 to 361,329** -- a drop of 22,256 students (-5.8%). This could signal declining birth rates, rising private school enrollment, or families delaying school entry.

```{r grade-shifts}
grade_trend <- enr %>%
  filter(is_state, subgroup == "total_enrollment",
         grade_level %in% c("PK", "K", "01", "05", "09", "12")) %>%
  select(end_year, grade_level, n_students) %>%
  pivot_wider(names_from = end_year, values_from = n_students) %>%
  mutate(
    change = `2024` - `2020`,
    pct_change = round(change / `2020` * 100, 1)
  )

stopifnot(nrow(grade_trend) > 0)
print(grade_trend)
```

Pre-K also dropped, while high school grades grew. The pipeline is narrowing at the entry point -- a trend that will ripple through the system for years.

---

## 8. 62% of Students Are Economically Disadvantaged

The share of **economically disadvantaged** students grew from 60.3% to 62.3% -- nearly two-thirds of all students in Texas public schools.

```{r econ-disadv}
econ_trend <- enr %>%
  filter(is_state, subgroup == "econ_disadv", grade_level == "TOTAL") %>%
  select(end_year, n_students, pct) %>%
  mutate(pct = round(pct * 100, 1))

stopifnot(nrow(econ_trend) > 0)
print(econ_trend)
```

```{r econ-plot}
print(econ_trend)

ggplot(econ_trend, aes(x = end_year, y = pct)) +
  geom_col(fill = "#56B4E9", width = 0.6) +
  geom_hline(yintercept = 50, linetype = "dashed", color = "gray50") +
  geom_text(aes(label = paste0(pct, "%")), vjust = -0.3, size = 3.5) +
  scale_y_continuous(limits = c(0, 70)) +
  labs(
    title = "Economically Disadvantaged Students",
    subtitle = "Percentage of total enrollment",
    x = "School Year (End)", y = "Percent"
  ) +
  theme_minimal()
```

Some large districts serve even higher concentrations of low-income students.

---

## 9. White Students Dropped Below 25%

**White enrollment** declined from 27.0% to 25.0% of total enrollment -- a 2 percentage point drop in just five years. Meanwhile, Hispanic (53.2%) and Asian (5.4%) shares continue to grow.

```{r demographic-shift}
demo_shift <- enr %>%
  filter(is_state, grade_level == "TOTAL",
         subgroup %in% c("white", "black", "hispanic", "asian", "multiracial")) %>%
  select(end_year, subgroup, pct) %>%
  mutate(pct = round(pct * 100, 1)) %>%
  pivot_wider(names_from = subgroup, values_from = pct)

stopifnot(nrow(demo_shift) > 0)
print(demo_shift)
```

```{r demo-plot}
demo_plot_data <- enr %>%
  filter(is_state, grade_level == "TOTAL",
         subgroup %in% c("white", "black", "hispanic", "asian")) %>%
  mutate(pct = pct * 100)

stopifnot(nrow(demo_plot_data) > 0)
print(demo_plot_data %>% select(end_year, subgroup, pct))

ggplot(demo_plot_data, aes(x = end_year, y = pct, color = subgroup)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2) +
  geom_vline(xintercept = 2020.5, linetype = "dotted", color = "gray50") +
  scale_color_manual(
    values = c("hispanic" = "#0072B2", "white" = "#E69F00",
               "black" = "#009E73", "asian" = "#CC79A7"),
    labels = c("Asian", "Black", "Hispanic", "White")
  ) +
  labs(
    title = "Texas Public School Demographics",
    x = "School Year (End)", y = "Percent of Enrollment",
    color = NULL
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

---

## 10. 439 Districts Now Have Hispanic Majorities

The number of districts where Hispanic students are the majority grew from 419 in 2020 to **439 in 2024** -- now 36.4% of all Texas districts.

```{r hispanic-majority}
hisp_majority <- enr %>%
  filter(is_district, subgroup == "hispanic", grade_level == "TOTAL") %>%
  mutate(majority = pct > 0.5) %>%
  group_by(end_year) %>%
  summarize(
    total_districts = n(),
    hispanic_majority = sum(majority),
    pct_majority = round(hispanic_majority / total_districts * 100, 1),
    .groups = "drop"
  )

stopifnot(nrow(hisp_majority) > 0)
print(hisp_majority)
```

20 districts flipped from non-majority to Hispanic majority between 2020 and 2024.

---

## 11. Houston ISD vs Dallas ISD: Two Giants, Two Trajectories

Houston ISD and Dallas ISD are the two largest traditional districts in Texas. Both lost students since 2020, but Houston's losses are nearly double Dallas's in absolute terms.

```{r houston-vs-dallas}
# Use district_id to include 2020 (which has NA district_name)
big_two_ids <- c("101912", "057905")  # Houston ISD, Dallas ISD

big_two <- enr %>%
  filter(is_district, subgroup == "total_enrollment", grade_level == "TOTAL",
         district_id %in% big_two_ids) %>%
  mutate(district_label = case_when(
    district_id == "101912" ~ "HOUSTON ISD",
    district_id == "057905" ~ "DALLAS ISD"
  )) %>%
  select(end_year, district_label, n_students) %>%
  pivot_wider(names_from = district_label, values_from = n_students) %>%
  arrange(end_year) %>%
  mutate(
    houston_change = `HOUSTON ISD` - lag(`HOUSTON ISD`),
    dallas_change = `DALLAS ISD` - lag(`DALLAS ISD`)
  )

stopifnot(nrow(big_two) > 0)
print(big_two)
```

```{r houston-vs-dallas-plot}
big_two_plot <- enr %>%
  filter(is_district, subgroup == "total_enrollment", grade_level == "TOTAL",
         district_id %in% big_two_ids) %>%
  mutate(district_label = case_when(
    district_id == "101912" ~ "HOUSTON ISD",
    district_id == "057905" ~ "DALLAS ISD"
  ))

stopifnot(nrow(big_two_plot) > 0)
print(big_two_plot %>% select(end_year, district_label, n_students))

ggplot(big_two_plot, aes(x = end_year, y = n_students, color = district_label)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  geom_vline(xintercept = 2020.5, linetype = "dotted", color = "gray50") +
  scale_y_continuous(labels = comma) +
  scale_color_manual(values = c("DALLAS ISD" = "#0072B2", "HOUSTON ISD" = "#CC0000")) +
  labs(
    title = "Houston ISD vs Dallas ISD Enrollment",
    subtitle = "Both declining, but Houston falling faster",
    x = "School Year (End)", y = "Total Students", color = NULL
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

Houston lost **25,706 students** (-12.3%) vs Dallas's **14,688** (-9.5%) between 2020 and 2024. Both are losing ground to suburban and charter competitors.

---

## 12. Nearly 1 in 7 Students Receives Special Education

In 2024, **764,858 students** -- 13.9% of all Texas public school students -- receive special education services. That means roughly 1 in 7 students has an identified disability requiring specialized support.

```{r sped-snapshot}
sped_2024 <- enr %>%
  filter(is_state, subgroup == "special_ed", grade_level == "TOTAL", end_year == 2024) %>%
  select(end_year, n_students, pct) %>%
  mutate(pct_display = round(pct * 100, 1))

stopifnot(nrow(sped_2024) > 0)
print(sped_2024 %>% select(end_year, n_students, pct_display))
```

Texas historically undercounted special education students due to a controversial 8.5% cap on special education identification. After the cap was lifted in 2018, identification rates have climbed significantly -- reaching 13.9% statewide.

---

## 13. Suburban Boomtowns Are Reshaping Texas Education

While urban districts lose students, fast-growing suburban districts are booming. **Hallsville ISD** nearly doubled (+85.7%), **Prosper ISD** grew 69.1%, and **Frisco ISD** added 3,980 students (+6.4%) between 2020 and 2024.

```{r suburban-growth}
# Top growing large districts (using district_id join for accurate 2020-2024 comparison)
suburban_growth <- losses %>%
  filter(n_2020 >= 5000) %>%
  arrange(desc(pct_change)) %>%
  select(district_name, n_2020, n_2024, change, pct_change) %>%
  head(10)

stopifnot(nrow(suburban_growth) > 0)
print(suburban_growth)
```

```{r frisco-trend-plot}
# Frisco ISD enrollment (2021-2024 by name, 2020 via district_id)
frisco_id <- "043905"
frisco_trend <- enr %>%
  filter(is_district, district_id == frisco_id,
         subgroup == "total_enrollment", grade_level == "TOTAL") %>%
  select(end_year, n_students) %>%
  arrange(end_year)

stopifnot(nrow(frisco_trend) > 0)
print(frisco_trend)

ggplot(frisco_trend, aes(x = end_year, y = n_students)) +
  geom_col(fill = "#56B4E9", width = 0.6) +
  geom_text(aes(label = comma(n_students)), vjust = -0.3, size = 3.5) +
  geom_vline(xintercept = 2020.5, linetype = "dotted", color = "gray50") +
  scale_y_continuous(labels = comma, limits = c(0, 75000)) +
  labs(
    title = "Frisco ISD Enrollment",
    subtitle = "Suburban growth in Collin County, 2020-2024",
    x = "School Year (End)", y = "Total Students"
  ) +
  theme_minimal()
```

The suburban boom reflects DFW's explosive population growth, with young families moving to master-planned communities in Prosper, Frisco, Princeton, and Forney.

---

## 14. Pre-K Enrollment Still Hasn't Recovered from COVID

Pre-K enrollment cratered by **21%** during COVID (from 248,413 to 196,560). By 2024 it has clawed back to 247,979 -- still **434 students short** of pre-pandemic levels.

```{r pk-recovery}
pk_trend <- enr %>%
  filter(is_state, subgroup == "total_enrollment", grade_level == "PK") %>%
  select(end_year, n_students) %>%
  arrange(end_year) %>%
  mutate(
    change = n_students - lag(n_students),
    pct_change = round(change / lag(n_students) * 100, 1),
    vs_2020 = n_students - first(n_students)
  )

stopifnot(nrow(pk_trend) > 0)
print(pk_trend)
```

```{r pk-recovery-plot}
print(pk_trend)

pk_2020_level <- pk_trend$n_students[pk_trend$end_year == 2020]

ggplot(pk_trend, aes(x = end_year, y = n_students)) +
  geom_line(linewidth = 1.2, color = "#D55E00") +
  geom_point(size = 3, color = "#D55E00") +
  geom_vline(xintercept = 2020.5, linetype = "dotted", color = "gray50") +
  geom_hline(yintercept = pk_2020_level, linetype = "dashed", color = "gray50") +
  annotate("text", x = 2022, y = pk_2020_level + 2500, label = "2020 level",
           color = "gray40", size = 3.5) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Pre-K Enrollment: The Slowest Recovery",
    subtitle = "Still below 2020 levels after four years",
    x = "School Year (End)", y = "Pre-K Students"
  ) +
  theme_minimal()
```

The slow Pre-K recovery matters because early childhood education is one of the highest-return investments in education.

---

## 15. Multiracial Students Are the Fastest-Growing Demographic

Students identifying as **Two or More Races** grew from 2.5% to 3.1% of enrollment between 2020 and 2024 -- a 25% increase in raw numbers (138,434 to 173,425). While still a small share, this is the fastest percentage growth of any demographic group.

```{r multiracial-growth}
multi_trend <- enr %>%
  filter(is_state, subgroup == "multiracial", grade_level == "TOTAL") %>%
  select(end_year, n_students, pct) %>%
  mutate(pct_display = round(pct * 100, 1))

stopifnot(nrow(multi_trend) > 0)
print(multi_trend %>% select(end_year, n_students, pct_display))
```

```{r multiracial-plot}
print(multi_trend %>% select(end_year, n_students, pct_display))

ggplot(multi_trend, aes(x = end_year, y = n_students)) +
  geom_line(linewidth = 1.2, color = "#882255") +
  geom_point(size = 3, color = "#882255") +
  geom_vline(xintercept = 2020.5, linetype = "dotted", color = "gray50") +
  geom_text(aes(label = comma(n_students)), vjust = -0.8, size = 3.5) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Multiracial Students: Fastest-Growing Group",
    subtitle = "Two or More Races enrollment, 2020-2024",
    x = "School Year (End)", y = "Students"
  ) +
  theme_minimal()
```

This trend mirrors national patterns and reflects both demographic change and evolving identity choices.

---

## Summary

These fifteen findings illustrate the rapid transformation of Texas public education:

| Finding | Change (2020-2024) |
|---------|-------------------|
| COVID enrollment drop | -120,133 students (2020-21) |
| LEP students | 20.3% to 24.4% (+4.1 pts) |
| Coppell ISD Asian | Now 56.7% Asian |
| Fort Worth ISD | -14.3% (-11,801 students) |
| IDEA Public Schools | +55.3% (+27,339 students) |
| Fort Bend ISD | No majority group (most diverse) |
| Kindergarten enrollment | -5.8% |
| Economically disadvantaged | 60.3% to 62.3% |
| White enrollment | 27.0% to 25.0% |
| Hispanic majority districts | 419 to 439 (+20) |
| Houston vs Dallas ISD | -25,706 vs -14,688 students |
| Special education | 13.9% of enrollment (764,858 students) |
| Suburban boomtowns | Hallsville +85.7%, Prosper +69.1% |
| Pre-K enrollment | Still below 2020 levels |
| Multiracial students | Fastest-growing group (+25%) |

The data tells a story of suburbanization, rising poverty, linguistic diversity, and demographic transformation that will shape Texas education for decades to come.

---

## Explore the Data

Use `txschooldata` to explore these patterns yourself:

```{r explore-example, eval=FALSE}
library(txschooldata)
library(dplyr)

# Fetch 2024 data
enr <- fetch_enr(2024, use_cache = TRUE)

# Your district's demographics
enr %>%
  filter(district_name == "YOUR DISTRICT NAME", grade_level == "TOTAL") %>%
  select(subgroup, n_students, pct)
```

## Session Info

```{r session-info}
sessionInfo()
```
