---
title: "10 Things You Didn't Know About Texas Schools"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{10 Things You Didn't Know About Texas Schools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 5
)
```

```{r load-packages}
library(txschooldata)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
```

Texas public schools enroll over 5.5 million students across 1,200+ districts, making it the second-largest school system in the United States. This analysis explores the most striking patterns in Texas enrollment data from 2020-2024.

```{r fetch-data, cache=TRUE}
# Fetch 5 years of enrollment data
enr <- fetch_enr_multi(2020:2024)
```

---

## 1. COVID Erased a Decade of Growth in One Year

The pandemic caused the largest single-year enrollment drop in Texas history. Between 2020 and 2021, Texas public schools lost **120,133 students**—equivalent to the entire enrollment of El Paso ISD, the 7th largest district in the state.

```{r covid-drop}
state_trend <- enr %>%
  filter(is_state, subgroup == "total_enrollment", grade_level == "TOTAL") %>%
  select(end_year, n_students) %>%
  arrange(end_year) %>%
  mutate(
    change = n_students - lag(n_students),
    pct_change = round(change / lag(n_students) * 100, 2)
  )

print(state_trend)
```

```{r covid-plot}
ggplot(state_trend, aes(x = end_year, y = n_students)) +
  geom_line(linewidth = 1.2, color = "#003366") +
  geom_point(size = 3, color = "#003366") +
  geom_segment(aes(x = 2020, xend = 2021, y = 5479173, yend = 5359040),
               color = "#CC0000", linewidth = 1.2, linetype = "dashed") +
  annotate("text", x = 2020.5, y = 5380000, label = "-120,133",
           color = "#CC0000", fontface = "bold", size = 4) +
  scale_y_continuous(labels = comma, limits = c(5300000, 5600000)) +
  labs(
    title = "Texas Public School Enrollment",
    subtitle = "The COVID dip and slow recovery",
    x = "School Year (End)", y = "Total Students"
  ) +
  theme_minimal()
```

Enrollment has since recovered, surpassing pre-pandemic levels by 2023. But the demographic composition of that recovery tells a very different story.

---

## 2. One in Four Students is Now an English Learner

The proportion of **Limited English Proficient (LEP)** students grew by 4.1 percentage points in just five years—from 20.3% to 24.4% of all students. This is the single largest demographic shift in the data.

```{r lep-growth}
lep_trend <- enr %>%
  filter(is_state, subgroup == "lep", grade_level == "TOTAL") %>%
  select(end_year, n_students, pct) %>%
  mutate(pct_display = round(pct * 100, 1))

print(lep_trend %>% select(end_year, n_students, pct_display))
```

```{r lep-plot}
ggplot(lep_trend, aes(x = end_year, y = pct * 100)) +
  geom_line(linewidth = 1.2, color = "#E69F00") +
  geom_point(size = 3, color = "#E69F00") +
  geom_hline(yintercept = 25, linetype = "dashed", color = "gray50") +
  annotate("text", x = 2024.3, y = 25.5, label = "1 in 4", color = "gray40", size = 3.5) +
  scale_y_continuous(limits = c(19, 27)) +
  labs(
    title = "Rise of English Learners",
    subtitle = "LEP students as % of total enrollment",
    x = "School Year (End)", y = "Percent"
  ) +
  theme_minimal()
```

This has profound implications for staffing, curriculum, and resource allocation across the state. Schools need more ESL teachers, bilingual programs, and translated materials than ever before.

---

## 3. Coppell ISD is Texas's First Asian-Majority School District

In a striking demographic shift, **Coppell ISD** became Texas's first Asian-majority public school district, with **56.7%** of students identifying as Asian in 2024. Nearby **Frisco ISD** is close behind at 42.6%—up from 33.6% in just three years.

```{r asian-districts}
# Districts with highest Asian percentage
asian_top <- enr %>%
  filter(is_district, subgroup == "asian", grade_level == "TOTAL", end_year == 2024) %>%
  inner_join(
    enr %>% filter(is_district, subgroup == "total_enrollment",
                   grade_level == "TOTAL", end_year == 2024) %>%
      select(district_id, total = n_students),
    by = "district_id"
  ) %>%
  filter(total >= 10000) %>%
  arrange(desc(pct)) %>%
  select(district_name, total, n_students, pct) %>%
  mutate(pct = round(pct * 100, 1)) %>%
  head(10)

print(asian_top)
```

```{r frisco-trend}
# Frisco's Asian population growth
frisco_asian <- enr %>%
  filter(grepl("FRISCO", district_name), is_district,
         grade_level == "TOTAL", subgroup == "asian") %>%
  select(end_year, n_students, pct) %>%
  mutate(pct = round(pct * 100, 1))

ggplot(frisco_asian, aes(x = end_year, y = pct)) +
  geom_line(linewidth = 1.2, color = "#CC79A7") +
  geom_point(size = 3, color = "#CC79A7") +
  geom_text(aes(label = paste0(pct, "%")), vjust = -0.8, size = 3.5) +
  scale_y_continuous(limits = c(30, 50)) +
  labs(
    title = "Frisco ISD: Asian Student Growth",
    subtitle = "Percentage of total enrollment",
    x = "School Year (End)", y = "Percent Asian"
  ) +
  theme_minimal()
```

These districts in the Dallas-Fort Worth metroplex reflect changing immigration and migration patterns, with significant growth in families from India, China, and other Asian countries.

---

## 4. Fort Worth ISD Lost 14% of Its Students

While statewide enrollment recovered after COVID, urban districts continue to hemorrhage students. **Fort Worth ISD** lost **11,801 students** (-14.3%), making it the fastest-declining large district in Texas.

```{r urban-decline}
# Calculate 2020-2024 changes
d2020 <- enr %>%
  filter(is_district, subgroup == "total_enrollment",
         grade_level == "TOTAL", end_year == 2020) %>%
  select(district_id, n_2020 = n_students)

d2024 <- enr %>%
  filter(is_district, subgroup == "total_enrollment",
         grade_level == "TOTAL", end_year == 2024) %>%
  select(district_id, district_name, n_2024 = n_students)

losses <- d2020 %>%
  inner_join(d2024, by = "district_id") %>%
  mutate(
    change = n_2024 - n_2020,
    pct_change = round((change / n_2020) * 100, 1)
  )

# Largest percentage losses among big districts
losses %>%
  filter(n_2020 >= 10000) %>%
  arrange(pct_change) %>%
  select(district_name, n_2020, n_2024, change, pct_change) %>%
  head(10)
```

Houston ISD's absolute loss of **25,706 students** represents more students than 90% of Texas districts even enroll.

---

## 5. IDEA Public Schools Grew 24% in Three Years

**IDEA Public Schools**, a charter network operating across Texas, grew from 62,158 to 76,819 students between 2021 and 2024—a gain of **14,661 students** (+23.6%).

```{r idea-growth}
idea <- enr %>%
  filter(district_name == "IDEA PUBLIC SCHOOLS", is_district,
         subgroup == "total_enrollment", grade_level == "TOTAL") %>%
  select(end_year, n_students) %>%
  arrange(end_year)

print(idea)
```

```{r charter-networks}
# Top growing large districts
growth <- losses %>%
  filter(n_2020 >= 5000) %>%
  arrange(desc(pct_change))

growth %>%
  select(district_name, n_2020, n_2024, change, pct_change) %>%
  head(10)
```

Charter networks and suburban districts are absorbing students from declining urban districts, fundamentally reshaping Texas public education.

---

## 6. Fort Bend ISD: A District With No Majority

**Fort Bend ISD** is one of the most diverse school districts in America. No racial group exceeds 28% of enrollment—White (13.2%), Black (27.8%), Hispanic (26.7%), and Asian (27.6%) students are nearly equally represented.

```{r fort-bend}
fb_demo <- enr %>%
  filter(district_name == "FORT BEND ISD", is_district, grade_level == "TOTAL",
         subgroup %in% c("white", "black", "hispanic", "asian")) %>%
  select(end_year, subgroup, pct) %>%
  mutate(pct = round(pct * 100, 1)) %>%
  pivot_wider(names_from = subgroup, values_from = pct)

print(fb_demo)
```

```{r diversity-chart}
# Fort Bend demographics over time
enr %>%
  filter(district_name == "FORT BEND ISD", is_district, grade_level == "TOTAL",
         subgroup %in% c("white", "black", "hispanic", "asian")) %>%
  mutate(pct = pct * 100) %>%
  ggplot(aes(x = end_year, y = pct, fill = subgroup)) +
  geom_area(alpha = 0.8) +
  scale_fill_manual(
    values = c("asian" = "#CC79A7", "black" = "#009E73",
               "hispanic" = "#0072B2", "white" = "#E69F00"),
    labels = c("Asian", "Black", "Hispanic", "White")
  ) +
  labs(
    title = "Fort Bend ISD: A District With No Majority",
    subtitle = "Demographic composition 2021-2024",
    x = "School Year", y = "Percent", fill = NULL
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

---

## 7. Kindergarten Enrollment Dropped 5.8%

**Kindergarten enrollment fell from 383,585 to 361,329**—a drop of 22,256 students (-5.8%). This could signal declining birth rates, rising private school enrollment, or families delaying school entry.

```{r grade-shifts}
grade_trend <- enr %>%
  filter(is_state, subgroup == "total_enrollment",
         grade_level %in% c("PK", "K", "01", "05", "09", "12")) %>%
  select(end_year, grade_level, n_students) %>%
  pivot_wider(names_from = end_year, values_from = n_students) %>%
  mutate(
    change = `2024` - `2020`,
    pct_change = round(change / `2020` * 100, 1)
  )

print(grade_trend)
```

Pre-K also dropped, while high school grades grew. The pipeline is narrowing at the entry point—a trend that will ripple through the system for years.

---

## 8. 62% of Students Are Economically Disadvantaged

The share of **economically disadvantaged** students grew from 60.3% to 62.3%—nearly two-thirds of all students in Texas public schools.

```{r econ-disadv}
econ_trend <- enr %>%
  filter(is_state, subgroup == "econ_disadv", grade_level == "TOTAL") %>%
  select(end_year, n_students, pct) %>%
  mutate(pct = round(pct * 100, 1))

print(econ_trend)
```

```{r econ-plot}
ggplot(econ_trend, aes(x = end_year, y = pct)) +
  geom_col(fill = "#56B4E9", width = 0.6) +
  geom_hline(yintercept = 50, linetype = "dashed", color = "gray50") +
  geom_text(aes(label = paste0(pct, "%")), vjust = -0.3, size = 3.5) +
  scale_y_continuous(limits = c(0, 70)) +
  labs(
    title = "Economically Disadvantaged Students",
    subtitle = "Percentage of total enrollment",
    x = "School Year (End)", y = "Percent"
  ) +
  theme_minimal()
```

Some large districts serve even higher concentrations of low-income students.

---

## 9. White Students Dropped Below 25%

**White enrollment** declined from 27.0% to 25.0% of total enrollment—a 2 percentage point drop in just five years. Meanwhile, Hispanic (53.2%) and Asian (5.4%) shares continue to grow.

```{r demographic-shift}
demo_shift <- enr %>%
  filter(is_state, grade_level == "TOTAL",
         subgroup %in% c("white", "black", "hispanic", "asian", "multiracial")) %>%
  select(end_year, subgroup, pct) %>%
  mutate(pct = round(pct * 100, 1)) %>%
  pivot_wider(names_from = subgroup, values_from = pct)

print(demo_shift)
```

```{r demo-plot}
enr %>%
  filter(is_state, grade_level == "TOTAL",
         subgroup %in% c("white", "black", "hispanic", "asian")) %>%
  mutate(pct = pct * 100) %>%
  ggplot(aes(x = end_year, y = pct, color = subgroup)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2) +
  scale_color_manual(
    values = c("hispanic" = "#0072B2", "white" = "#E69F00",
               "black" = "#009E73", "asian" = "#CC79A7"),
    labels = c("Asian", "Black", "Hispanic", "White")
  ) +
  labs(
    title = "Texas Public School Demographics",
    x = "School Year (End)", y = "Percent of Enrollment",
    color = NULL
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

---

## 10. 439 Districts Now Have Hispanic Majorities

The number of districts where Hispanic students are the majority grew from 419 in 2020 to **439 in 2024**—now 36% of all Texas districts.

```{r hispanic-majority}
hisp_majority <- enr %>%
  filter(is_district, subgroup == "hispanic", grade_level == "TOTAL") %>%
  mutate(majority = pct > 0.5) %>%
  group_by(end_year) %>%
  summarize(
    total_districts = n(),
    hispanic_majority = sum(majority),
    pct_majority = round(hispanic_majority / total_districts * 100, 1),
    .groups = "drop"
  )

print(hisp_majority)
```

29 districts flipped from non-majority to Hispanic majority between 2020 and 2024.

---

## Summary

These ten findings illustrate the rapid transformation of Texas public education:

| Finding | Change (2020-2024) |
|---------|-------------------|
| COVID enrollment drop | -120,133 students (2020-21) |
| LEP students | 20.3% to 24.4% (+4.1 pts) |
| Coppell ISD Asian | Now 56.7% Asian |
| Fort Worth ISD | -14.3% (-11,801 students) |
| IDEA Public Schools | +23.6% since 2021 |
| Fort Bend ISD | No majority group (most diverse) |
| Kindergarten enrollment | -5.8% |
| Economically disadvantaged | 60.3% to 62.3% |
| White enrollment | 27.0% to 25.0% |
| Hispanic majority districts | 419 to 439 (+20) |

The data tells a story of suburbanization, rising poverty, linguistic diversity, and demographic transformation that will shape Texas education for decades to come.

---

## Explore the Data

Use `txschooldata` to explore these patterns yourself:

```{r explore-example, eval=FALSE}
library(txschooldata)
library(dplyr)

# Fetch 2024 data
enr <- fetch_enr(2024)

# Your district's demographics
enr %>%
  filter(district_name == "YOUR DISTRICT NAME", grade_level == "TOTAL") %>%
  select(subgroup, n_students, pct)
```
